### 2017-01-05 doc/manuscript.Rmd
Added basic skeleton of discussion (with refs)

Title:

@@

Limitations:
- While GEE is well suited to longitudinal datasets.. threre is a possibility
given teh complex biology that the temporal associations can not be adequately
analyzed using current methodologies. IN particular, an important issue si taht
we did not quantify TAGFA at each of the three visits and so can not directed
infer changes in TAGFA with changes in IR. However, given potential feedback
between IR and TAGFA, this may in fact be a strength for this particular
association given we are trying to predict how a FA at one time point can
predict future values.
- While PLS is a well-established technique for constructing predictive models
of high dimensionality data strcutures, it can not adjust for potential
confounders and other effect modifiers. Likewise, PLS is not able to handle
longitudinal data so only the baseline visit was used in the PLS analysis.
However, given the limitation, there results are concordant between GEE and PLS
models.
- Our population was predocminiately female and of euro ancestry, who were at 
riskj for diabetes at study enrollment, and as such our results may not be
generalizable to other populations.
- Given the complex biology underlying the variables of interest, there many
some residual confounding we are unaware of or have not been able to adjust for.
However, we have taken extensive precautions for trying to ID confounders
through the DAG models and QIC computations.

@@

- previous research has shown that CHO intake increases DNL {{ cite }}. In
particular, DNL from CHO sources increases especially the 14-16 chain TAGFA as
well as the 18 chain TAGFA. Prev lit has also shown that higher DNL can increase
risk for DM {{ cite }} and that higher refined CHO can increase risk of DM {{
cite }}. Our study found that the 14-16 TAGFA all strongly predicted lower
insulin sensitivity, reinforcing this finding. While these FA also had a
significant association with beta-cell function, the results were not as strong
as with the IR. {{ why is it that DNL increases risk? }}
- TAGFA that tended to positively associated with IR and BCF also tended to be
the very long chain FA such as eicosapentaenoic acid (20:5n-3) and
docosahexaenoic acid (22:6n-3), which have potent anti-inflammatory properties
{{ cite }}. Biologically {{ why these FA and how help with IR }}
- {{ compare models with and without NEFA }} NEFA have been shown contribute
substantially to TAGFA composition and concentration. Adjusting for ...
- Dietary FA have been shown to contribute approx 25% of the newly synth
VLDL-TAG {{ cite }}. This suggests that our results may not entirely reflect
diet, but also adipose stores. {{ discuss results of NEFA paper }} However, in
an analysis of the NEFA pool {{ cite }}, we found that total NEFA, but not the
specific composition associated with beta-cell function. There may be something
unique about the specific composition of the TAGFA fraction that may allow it to
enter IR tissues more easily than NEFA. It may be that VLDL receptors allow FA
to transition to IR tissue compared to NEFA bound to albumin {{ confirm and cite
}}.

Previous lit and limits:

-

@@

[@Rhee2011a] fasting TAG status better as discriminating between DM case vs
control than 2hr-OGTT TAG

[@Kotronen2009a] (this wasn't studied, but discussed) higher hepatic DNL may
increase amount of 'shorter' chain FA (16:0, 14:0, etc.) into VLDL and hence
into circulation. So higher of these FA may indicate higher DNL -> problems with
liver function -> maybe hepatic IR -> cyclical?

Discuss my PL and CE paper as well. Maybe also the NEFA one if it is published
by then...

The four FA that cluster together on the PLS also appear to be the same FA that
make up most of DNL, especially in context of higher carb/refined carb in diet
[@Wilke2009a].

[@Lee2015a] 16:1n-7 may be good marker of liver DNL.

>>>>

Title: Fatty acids involved in de novo lipogenesis within triacylglycerol fraction strongly predict lower insulin sensitivity.

@@

@@

- previous research has shown that CHO intake increases DNL {{ cite }}. In
particular, DNL from CHO sources increases especially the 14-16 chain TAGFA as
well as the 18 chain TAGFA. Prev lit has also shown that higher DNL can increase
risk for DM {{ cite }} and that higher refined CHO can increase risk of DM {{
cite }}. Our study found that the 14-16 TAGFA all strongly predicted lower
insulin sensitivity, reinforcing this finding. While these FA also had a
significant association with beta-cell function, the results were not as strong
as with the IR. {{ why is it that DNL increases risk? }}
- TAGFA that tended to positively associated with IR and BCF also tended to be
the very long chain FA such as eicosapentaenoic acid (20:5n-3) and
docosahexaenoic acid (22:6n-3), which have potent anti-inflammatory properties
{{ cite }}. Biologically {{ why these FA and how help with IR }}
- {{ compare models with and without NEFA }} NEFA have been shown contribute
substantially to TAGFA composition and concentration. Adjusting for ...
- Dietary FA have been shown to contribute approx 25% of the newly synth
VLDL-TAG {{ cite }}. This suggests that our results may not entirely reflect
diet, but also adipose stores. {{ discuss results of NEFA paper }} However, in
an analysis of the NEFA pool {{ cite }}, we found that total NEFA, but not the
specific composition associated with beta-cell function. There may be something
unique about the specific composition of the TAGFA fraction that may allow it to
enter IR tissues more easily than NEFA. It may be that VLDL receptors allow FA
to transition to IR tissue compared to NEFA bound to albumin {{ confirm and cite
}}.
- Challenge longitudinally since muscle IR can promote DNL + TAG Flannery2012a
- Or that post-prandial TAG is not suppressed in IR Sondergaard2012a
- Differences with 
- {{ confirm }} greater flux of saturated fatty acids into tissues may encourage
lipotoxic environment {{ cite }}

Limitations:
- While GEE is well suited to longitudinal datasets.. threre is a possibility
given teh complex biology that the temporal associations can not be adequately
analyzed using current methodologies. IN particular, an important issue si taht
we did not quantify TAGFA at each of the three visits and so can not directed
infer changes in TAGFA with changes in IR. However, given potential feedback
between IR and TAGFA, this may in fact be a strength for this particular
association given we are trying to predict how a FA at one time point can
predict future values.
- While PLS is a well-established technique for constructing predictive models
of high dimensionality data strcutures, it can not adjust for potential
confounders and other effect modifiers. Likewise, PLS is not able to handle
longitudinal data so only the baseline visit was used in the PLS analysis.
However, given the limitation, there results are concordant between GEE and PLS
models.
- Our population was predocminiately female and of euro ancestry, who were at 
riskj for diabetes at study enrollment, and as such our results may not be
generalizable to other populations.
- Given the complex biology underlying the variables of interest, there many
some residual confounding we are unaware of or have not been able to adjust for.
However, we have taken extensive precautions for trying to ID confounders
through the DAG models and QIC computations.
- TAGFA may fluctuate daily, but we had participants come in fasted in the morning

Previous lit and limits:
- A few epidemiological studies have examined the role of the TAGFA composition
on insulin resistance and diabetes risk. 
- A weight loss intervention trial showed that changes in the TAGFA composition
of 19 participants accompanied changes in the weight and insulin sensitivity over
the 33 week period [@Schwab2008a]. Specifically, the TAGFA 14:0, 14:1, 16:0,
16:1, 18:0, 18:3n-3, and 18:3n-6 were all lower after weight loss and correlated
with improved insulin sensitivity.
- In a short (24hr) feeding trial of 24 subjects [@Lee2015a], palmitoleic acid
(16:1n-7) was shown to be strongly correlated with isotopically measured DNL and
was 45% higher in individuals with a greater amount of hepatic fat. However, in
that study only the TAGFA 18:0 was correlated with HOMA-IR, likely due to the
small sample size.
- To our knowledge, there have been two prospective studies that had a similar
research objective as our cohort. One was a nested case-control cohort of the 
Framingham offspring study [@Rhee2015a], which found that TAGFA with a lower
carbon chain and less double bonds (e.g. 14:0, 16:0) had higher OR for diabetes
after 12-years while higher carbon chain and more double bond TAGFA had a lower
OR for diabetes. This pattern of TAGFA with diabetes risk remained the same when
they examined HOMA-IR cross-sectionally at the baseline visit. The other study 
was a prospective cohort of males in Finland [@Lankinen2015a], of which they had
TAGFA data on 831 participants after 6-years from the baseline visit. This
cohort has similar outcomes as PROMISE, however, OGTT data was only available at
the 6-year visit. A beta-cell function measure was computed, but only included 
calculations up to 30 minutes of the OGTT. As with our results, they found a
larger effect size for insulin sensitivity compared to the beta-cell function.
The specific TAGFA that associated with the metabolic measures were similar 
between studies, except they did not find 
- The primary limit to these studies is the mainly small sample size and short
study duration. Two studies had larger sample sizes of 831 [@Lankinen2015a] and 
`r 189+189` [@Rhee2011a] and both were prospective cohorts. In addition, the
Framingham offspring study [@Rhee2011a] did not have data on beta-cell function 
while the Finnish cohort [@Lankinen2015a] used {{ how do I say this politely Tony? 
They're methods and article was ... questionable... }} simple statistical 
techniques to analysis the longitudinal data, which do not adequately take into
consideration auto-correlation of longitudinal values and the high dimensionality
of the TAGFA composition data structure.

- Refs:
    - Lankinen2015a (poor study, but is unfortunately relevant...) (god... I
    really don't want to cite/talk about this...)
    - Lee2015a (16:1n7 as marker of DNL) (n=24, feeding trial, 16:1n7 was 45%
    higher in high liver fat group, direct dnl (isotope) was 2.5 fold higher in
    high liver fat, 16:1n7 was corr (0.56) with direct DNL, TAG 18:0 corr (.50)
    with HOMA)
    - Zulyniak2012a (only vaccenic acid in TAG) (n=20, TLC-GC, HOMA-IR/HOMA-B,
    two groups: normogly and hypergly. results in general showed no diff between
    groups, only cis-vaccenic was higher in hyper, 'trend' with palmitic, did
    not present results on whole fraction for IR)
    - Rhee2011a (good paper on TAGFA composition) (Offspring of Framingham...,
    nested case-control, n=189 case, n=189 control, MS/MS, OGTT, differences in
    fasting TAGFA were more apparent than at 2-hr ogtt, more lower chain/lower
    desat in cases + higher OR, while higher chain/desat had lower OR, lower
    chain/desat pos corr with HOMA-IR, MS/MS (tandem) showed that 12-18 SFA and
    MUFA inc risk of diabetes)
    - Kotronen2009a (TAGFA composition) (n=16, 18-60, mostly female, lipidomics,
    GC, HOMA-IR, total esterified (of which TAGFA makes up the majority) 16,
    161n7, 181n9 pos corr with homa, 182n6 neg with homa)
    - Schwab2008a (TAGFA composition and insulin sensitivity) (n=75 (but only
    n=19 for lipidomics), overweight or obese, weght loss intervention (33
    weeks), weight, FSIGT, GC, UPLC-MS, after weight loss 14, 141, 16, 161, 18,
    183n3, 183n6 were lower, lower TAGFA that contained 14, 14:1, 16, 161, 18,
    181 were correlated with improved Si)

[@Rhee2011a] fasting TAG status better as discriminating between DM case vs
control than 2hr-OGTT TAG
[@Kotronen2009a] (this wasn't studied, but discussed) higher hepatic DNL may
increase amount of 'shorter' chain FA (16:0, 14:0, etc.) into VLDL and hence
into circulation. So higher of these FA may indicate higher DNL -> problems with
liver function -> maybe hepatic IR -> cyclical?
[@Lee2015a] 16:1n-7 may be good marker of liver DNL.

Wilke2009a - contributions to DNL (e.g. 70% 16:0 to newly synth)
The four FA that cluster together on the PLS also appear to be the same FA that
make up most of DNL, especially in context of higher carb/refined carb in diet
[@Wilke2009a].

@@
